---
title: "README"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

One of the best packages for pulling census data is `tidycensus`.  Other useful packages include `here` for file directory support, `tidyverse` for data manipulation, and `ggplot` for graphs.

```{r packages}
library(here)
library(tidyverse)

library(tidycensus)
options(tigris_use_cache = TRUE)

library(scales)
library(leaflet)
library(cowplot)
library(units)
library(sf)
```

## Setup

The following code sets up a blank theme so the maps look better and pulls in the county list to limit the data collection to western North Carolina.

```{r setup2}
## Null Theme
theme_null <- function() {
  theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.line = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_line(colour = 'transparent'),
        panel.grid.minor = element_blank())
}

counties <- read.csv(here::here("census", "Access", "R", "nc_counties.csv"))
counties <- counties %>% filter(WNC == 1)
```

## Census API key

You'll need to get your own API key.  This can be obtained [here](https://api.census.gov/data/key_signup.html).

```{r census_api_key}
# census_api_key("KEY GOES HERE", install = TRUE)
# Explain this better
```

## See variables

Here are all the available variables.

```{r variables}
acs_vars <- load_variables(2016, "acs5", cache = TRUE)

get_vars <- data.frame( name = c(
  "B01003_001E", # Total Population
  
  "B02001_001E", # Race/Ethnicity
  "B03002_003E",
  "B03002_004E",
  "B03002_005E",
  "B03002_006E",
  "B03002_007E",
  "B03002_008E",
  "B03002_010E",
  "B03002_011E",
  "B03002_012E"))

get_vars <- get_vars %>%
  left_join(acs_vars, by = "name")

# Clean up
get_vars$label <- gsub("Estimate|!|,|\\$", "", get_vars$label)
get_vars$label <- ifelse(get_vars$label != "Total", gsub("Total", "", get_vars$label), get_vars$label)
get_vars$label <- gsub("\\(([^\\)]+)\\)||Civilian labor force", "", get_vars$label)
get_vars$concept <- gsub("\\(([^\\)]+)\\)", "", get_vars$concept)
get_vars$label <- tolower(get_vars$label)
get_vars$concept <- tolower(get_vars$concept)
get_vars$label <- gsub(" ", "_", get_vars$label)
get_vars$concept <- gsub(" ", "_", get_vars$concept)
get_vars$label <- paste(get_vars$concept, get_vars$label, sep = "_")
get_vars$concept <- NULL
```

```{r population}
## ZCTA
# Close to Zip Code
zcta_county_match <- read.table(here::here("census", "Access", "R", "zcta_county_rel_10.txt"), # Get this file from census website
                                header = TRUE, sep = ",")

zcta_county_match <- zcta_county_match[zcta_county_match$STATE == 37 &
                                         zcta_county_match$COUNTY %in% counties$COUNTYFP, ]

start_time <- Sys.time()
wnc_zcta_data <- get_acs(geography = "zip code tabulation area",
                               variables = get_vars$name,
                               geometry = TRUE,
                               output = "wide")
Sys.time() - start_time
# 2 mins

wnc_zcta_data <- wnc_zcta_data[wnc_zcta_data$GEOID %in% zcta_county_match$ZCTA5,
                                           names(wnc_zcta_data) %in% c("NAME", "GEOID", "geometry", get_vars$name)]

## Counties
wnc_county_data <- reduce(
  map(counties$COUNTY, function(x) {
    get_acs(geography = "county", variables = get_vars$name,
            state = "NC", county = x, geometry = TRUE, output = "wide")
  }),
  rbind
)

wnc_county_data <- wnc_county_data[, names(wnc_county_data) %in% c("NAME", "GEOID", "geometry", get_vars$name)]

### Investigate

# Base Maps
ggplot() +
  geom_sf(data = wnc_county_data, col = "darkblue", fill = NA, lwd = 1.5) +
  geom_sf(data = wnc_zcta_data, col = "darkred", fill = NA) +
  ggtitle("Base Map") +
  theme_null()

# leaflet() %>%
#   addProviderTiles(provider = "CartoDB.Positron") %>%
#   addPolygons(data = wnc_county_data,
#               popup = ~ str_extract(NAME, "^([^,]*)"),
#               stroke = TRUE,
#               smoothFactor = .5,
#               fillColor = "white",
#               color = "darkblue",
#               weight = 2) %>%
#   addPolygons(data = wnc_zcta_data,
#               popup = ~ str_extract(NAME, "^([^,]*)"),
#               stroke = TRUE,
#               smoothFactor = .5,
#               fillColor = "white",
#               color = "darkred",
#               weight = 1)

## Total Population
# Add in Density
wnc_county_data <- wnc_county_data %>%
  mutate(area = as.numeric(set_units(st_area(geometry), mi^2)),
  density = as.numeric(B01003_001E / set_units(st_area(geometry), mi^2)))

wnc_zcta_data <- wnc_zcta_data %>%
  mutate(area = as.numeric(set_units(st_area(geometry), mi^2)),
         density = as.numeric(B01003_001E / set_units(st_area(geometry), mi^2)))

# County
county_1 <- ggplot() +
  geom_sf(data = wnc_county_data, aes(col = B01003_001E, fill = B01003_001E)) +
  scale_color_continuous(name = "Total Population",
                         labels = comma) +
  scale_fill_continuous(name = "Total Population",
                        labels = comma) +
  theme(legend.position = c(.25, .75)) +
  theme_null()

county_2 <- ggplot() +
  geom_sf(data = wnc_county_data, aes(col = density, fill = density)) +
  scale_color_continuous(name = "Population Density",
                         labels = comma) +
  scale_fill_continuous(name = "Population Density",
                        labels = comma) +
  theme(legend.position = c(.25, .75)) +
  theme_null()

plot_grid(county_1, county_2)

zcta_1 <- ggplot() +
  geom_sf(data = wnc_zcta_data, aes(col = B01003_001E, fill = B01003_001E)) +
  scale_color_continuous(name = "Total Population",
                         labels = comma) +
  scale_fill_continuous(name = "Total Population",
                        labels = comma) +
  theme(legend.position = c(.25, .75)) +
  theme_null()

zcta_2 <- ggplot() +
  geom_sf(data = wnc_zcta_data, aes(col = density, fill = density)) +
  scale_color_continuous(name = "Population Density",                         
                         labels = comma) +  
  scale_fill_continuous(name = "Population Density",
                        labels = comma) +  
  theme(legend.position = c(.25, .75)) +  
  theme_null()

plot_grid(zcta_1, zcta_2)
```
