---
title: "README"
output: md_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

One of the best packages for pulling census data is `tidycensus`.  More information about this package can be found at the creater's (site)[https://walkerke.github.io/tidycensus/].  This tutorial will cover the basics and focus on western North Carolina.  Other useful packages include `here` for file directory support, `tidyverse` for data manipulation, and `ggplot2` for graphs.

```{r packages, message = FALSE}
library(here)
library(ggplot2)
library(tidyverse)

library(tidycensus)
options(tigris_use_cache = TRUE)

library(scales)
library(leaflet)
library(cowplot)
library(units)
library(sf)
```

## Setup

The following code sets up a blank theme so the maps look better and pulls in the county list to limit the data collection to western North Carolina.  You can view the list of in the csv listed above this tutorial.

```{r setup2}
## Null Theme
theme_null <- function() {
  theme(axis.title=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.line = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank(),
        panel.grid.major = element_line(colour = 'transparent'),
        panel.grid.minor = element_blank())
}

counties <- read.csv(here::here("census", "Access", "R", "nc_counties.csv"))
counties <- counties %>% filter(WNC == 1)
```

## Census API key

You'll need to get your own API key.  This can be obtained [here](https://api.census.gov/data/key_signup.html).

```{r census_api_key}
# census_api_key("KEY GOES HERE", install = TRUE)
# Explain this better
```

## See variables

Here are all the available variables.

```{r variables}
acs_vars <- load_variables(2016, "acs5", cache = TRUE)

get_vars <- data.frame( name = c(
  "B01003_001E", # Total Population
  
  "B02001_001E", # Race/Ethnicity
  "B03002_003E",
  "B03002_004E",
  "B03002_005E",
  "B03002_006E",
  "B03002_007E",
  "B03002_008E",
  "B03002_010E",
  "B03002_011E",
  "B03002_012E"))

head(acs_vars)
```

## Pull data

The following code pulls the data from the census.  For North Carolina, the the STATE value is 37.  The zcta_county_rel_10.txt file came from the census website [here](https://www.census.gov/geo/maps-data/data/zcta_rel_layout.html).  

ZIP Code Tabulation Areas (ZCTAs) are similar to ZIP Codes but useful for census data.  More information about zip code tabulation area can be found [here](https://www.census.gov/geo/reference/zctas.html).

```{r zcta, message = FALSE}
## ZCTA
# Close to Zip Code
zcta_county_match <- read.table(here::here("census", "Access", "R", "zcta_county_rel_10.txt"), # Get this file from census website
                                header = TRUE, sep = ",")

zcta_county_match <- zcta_county_match[zcta_county_match$STATE == 37 &
                                         zcta_county_match$COUNTY %in% counties$COUNTYFP, ]

wnc_zcta_data <- get_acs(geography = "zip code tabulation area",
                               variables = get_vars$name,
                               geometry = TRUE,
                               output = "wide")
wnc_zcta_data <- wnc_zcta_data[wnc_zcta_data$GEOID %in% zcta_county_match$ZCTA5,
                                           names(wnc_zcta_data) %in% c("NAME", "GEOID", "geometry",
                                                                       as.character(get_vars$name))]
```

```{r counties}

## Counties
wnc_county_data <- reduce(
  map(counties$COUNTY, function(x) {
    get_acs(geography = "county", variables = get_vars$name,
            state = "NC", county = x, geometry = TRUE, output = "wide")
  }),
  rbind
)

wnc_county_data <- wnc_county_data[, names(wnc_county_data) %in% c("NAME", "GEOID", "geometry", as.character(get_vars$name))]
```

This map shows how the ZCTA and counties overlay each other.

```{r base_map}
### Investigate

# Base Maps
ggplot() +
  geom_sf(data = wnc_county_data, col = "darkblue", fill = NA, lwd = 1.5) +
  geom_sf(data = wnc_zcta_data, col = "darkred", fill = NA) +
  ggtitle("Base Map") +
  theme_null() +
  labs(caption = "Counties are thick dark blue lines.\nZCTA are thin red lines.")
```

This map shows population total and density for western NC ZTCA's.

```{r zcta_pop}
## Total Population
wnc_zcta_data <- wnc_zcta_data %>%
  mutate(area = as.numeric(set_units(st_area(geometry), mi^2)),
         density = as.numeric(B01003_001E / set_units(st_area(geometry), mi^2)))

zcta_1 <- ggplot() +
  geom_sf(data = wnc_zcta_data, aes(col = B01003_001E, fill = B01003_001E)) +
  scale_color_continuous(name = "Total Population",
                         labels = comma) +
  scale_fill_continuous(name = "Total Population",
                        labels = comma) +
  theme(legend.position = c(0, 1)) +
  theme_null()

zcta_2 <- ggplot() +
  geom_sf(data = wnc_zcta_data, aes(col = density, fill = density)) +
  scale_color_continuous(name = "Population Density",                         
                         labels = comma) +  
  scale_fill_continuous(name = "Population Density",
                        labels = comma) +  
  theme(legend.position = c(0, 1)) +  
  theme_null()

plot_grid(zcta_1, zcta_2)
```

This map shows population total and density for western NC counties.

```{r county_pop}
# Add in Density
wnc_county_data <- wnc_county_data %>%
  mutate(area = as.numeric(set_units(st_area(geometry), mi^2)),
  density = as.numeric(B01003_001E / set_units(st_area(geometry), mi^2)))

# County
county_1 <- ggplot() +
  geom_sf(data = wnc_county_data, aes(col = B01003_001E, fill = B01003_001E)) +
  scale_color_continuous(name = "Total Population",
                         labels = comma) +
  scale_fill_continuous(name = "Total Population",
                        labels = comma) +
  theme(legend.position = c(0, 1)) +
  theme_null()

county_2 <- ggplot() +
  geom_sf(data = wnc_county_data, aes(col = density, fill = density)) +
  scale_color_continuous(name = "Population Density",
                         labels = comma) +
  scale_fill_continuous(name = "Population Density",
                        labels = comma) +
  theme(legend.position = c(0, 1)) +
  theme_null()

plot_grid(county_1, county_2)
```

The follow code pulls data for census block for Buncombe County.

```{r buncombe}
buncombe_cb_data <- get_acs(geography = "block group",
                         variables = get_vars$name,
                         geometry = TRUE,
                         output = "wide",
                         state = 37,
                         county = "Buncombe")

buncombe_cb_data <- buncombe_cb_data %>%
  mutate(area = as.numeric(set_units(st_area(geometry), mi^2)),
         density = as.numeric(B01003_001E / set_units(st_area(geometry), mi^2)))

cb_1 <- ggplot() +
  geom_sf(data = buncombe_cb_data, aes(col = B01003_001E, fill = B01003_001E)) +
  scale_color_continuous(name = "Total Population",
                         labels = comma) +
  scale_fill_continuous(name = "Total Population",
                        labels = comma) +
  theme(legend.position = c(0, 1)) +
  theme_null()

cb_2 <- ggplot() +
  geom_sf(data = buncombe_cb_data, aes(col = density, fill = density)) +
  scale_color_continuous(name = "Population Density",
                         labels = comma) +
  scale_fill_continuous(name = "Population Density",
                        labels = comma) +
  theme(legend.position = c(0, 1)) +
  theme_null()

plot_grid(cb_1, cb_2)
```
